Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Recovery, and Limits to Cosmology

### 2017-01-24 paper1.tex
more edits

\subsection{The Buzzard Mock Catalogs}\label{sec: Buzzard}
The Buzzard mock galaxy catalogs are designed to mimic the sky geometry of DES and provide a light cone covering 398.49 \degsq\ between $4^h< RA < 6^h$ and $-61\degree < DEC < -41\degree$. They are derived from an in-house n-body cosmological simulation. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\sc 2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\sc LGadget-2} (a version of {\sc Gadget-2}; \citealt{Springel2005}). The Buzzard catalogs adopt the following cosmological parameters: $\Omega_m = 0.286$, $\Omega_\Lambda = 0.714$, $H_0 = 70$ \kms \mpc, $\sigma_8 = 0.82$, and $n_s = 0.96$. The DM halos are identified using the {\sc ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses.

@@

% \begin{figure}
% 	\includegraphics[width=\textwidth]{figures1/hmf.pdf}
% 	\caption{The cumulative MF (from \citealt{Tinker2008}) of halos above $M_{200c}$ at $z=0.1$}
% 	\label{fig: hmf}
% \end{figure}

@@

\subsection{Conditional {\rm[\ion{O}{II}]} Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS Data Release 12 \citep{Alam2015} catalogs to assign [\ion{O}{II}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{II}] line flux signal-to-noise of at least five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies, and the following SQL query selects the appropriate galaxies:
\begin{lstlisting}[language=SQL]
SELECT GAL.objid, 
       GAL.u, 
       GAL.err_u, 
       GAL.g, 
       GAl.err_g, 
       GAL.r, 
       GAL.err_r, 
       GAL.i, 
       GAl.err_i, 
       GAL.z, 
       GAL.err_z, 
       g.oii_3726_flux, 
       g.oii_3726_flux_err, 
       s.z 
FROM   galaxy AS GAL 
       JOIN specobj AS s 
         ON s.bestobjid = GAL.objid 
       JOIN galspecline AS g 
         ON g.specobjid = s.specobjid 
WHERE  oii_3726_flux > oii_3726_flux_err * 5 
       AND oii_3726_flux_err > 0 
       AND s.class = 'GALAXY' 
       AND s.zwarning = 0 
\end{lstlisting}

@@

The Hobby Eberly Telescope Dark Energy eXperiment (HETDEX; \citealt{Hill2008}) will observe high-redshift large scale structures using cutting edge wide-field integral field unit (IFU) spectrographs. Designed to probe the evolution of the dark energy equation of state etched onto high redshift ($z>2$) galaxies by the Baryon Acoustic Oscillations (BAO) \citep{Eisenstein2005}, the survey will observe two fields for a total of 420 \degsq\ (300 \degsq, Spring field and 120 \degsq, Fall field). Tuned to find \lya\ emitting (LAE) galaxies at $1.9<z<3.5$, HETDEX expects to find 800,000 LAEs, and more than one million [\ion{O}{II}] emitting galaxies at $z<0.5$ masquerading as high-redshift galaxies \citep{Acquaviva2014}.

While a large portion of the $\sim10^6$ interloping lower redshift galaxies will be field (not associated with a bound structure) galaxies, the large area covered by HETDEX is expected to contain as many as 50 Virgo-sized (halo mass $>10^{15}$ \msol) clusters at $z<0.5$. The near-complete spectroscopic coverage allows an unprecedentedly detailed look at a very large number of clusters ranging from group scales to the very massive. In addition to the recovery of accurate dynamical masses, detailed investigations of the dynamical state of the clusters is possible.

It is unclear how a blind spectroscopic survey with an IFU will effect the recovery of galaxy cluster dynamical properties. Unlike many previous large cluster surveys \citeeg{Milvang-Jensen2008, Robotham2011, Sifon2015} which use multi-object spectrographs, the Visible Integral-Field Replicable Unit Spectrograph (VIRUS; \citealt{Hill2012}) used by HETDEX samples the sky in a uniform but sparse way which could exclude member galaxies which would otherwise be included. Secondly, it is not straightforward to use spectroscopic redshifts predominately from emission-line galaxies to interpret the kinematic and dynamical states of the clusters.

HETDEX will measure the redshifts of $8 \times 10^5$ LAE galaxies between $1.9 < z < 3.5$ using a collection of 78 wide-field IFU spectrographs covering the wavelength region $3500 - 5500$~\AAA\ at R $\sim$ 750 \citep{Hill2008}. The primary science goal of these observations is to provide $<1\%$ accuracy measurements of the Hubble expansion parameter and the angular diameter distance at $z\sim2$. This result will provide significant constraints on the evolution of the dark energy equation of state which is both competitive with, and independent of, constraints derived from observations of the \lya forest.

The entire HETDEX survey will cover 420 \degsq\ with a 1/4.5 filling factor over two fields: a $\sim 300$ \degsq\ northern field, and a $\sim 140$ \degsq\ equatorial region. The spectral coverage allows for the detection of [\ion{O}{II}] ($\lambda\lambda 3727,3729$) emitters to $z\sim 0.5$ and Ca H ($\lambda 3968.5$) and K ($\lambda 3933.7$) absorption features to $z\sim 0.4$. The $10 \sigma$ detection threshold for emission spectral features will be $3.5\times10^{-17}$ \ergscm\ at 5000 \AAA, or equivalently, $g = 21.9$ mag for continuum objects.

@@

>>>>

The existing SDSS imaging will play an important role as the main HETDEX overlaps with the SDSS.

\subsection{The Buzzard Mock Catalogs}\label{sec: Buzzard}
The Buzzard mock galaxy catalogs are designed to mimic the sky geometry of DES and provide a light cone covering 398.49 \degsq\ between $4^h< RA < 6^h$ and $-61\degree < DEC < -41\degree$. They are derived from an in-house n-body cosmological simulation. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\sc 2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\sc LGadget-2} (a version of {\sc Gadget-2}; \citealt{Springel2005}). The simulation adopts the following cosmological parameters: $\Omega_m = 0.286$, $\Omega_\Lambda = 0.714$, $H_0 = 70$ \kms \mpc, $\sigma_8 = 0.82$, and $n_s = 0.96$. The DM halos are identified using the {\sc ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses.

@@

% \begin{figure}
% 	\includegraphics[width=\columnwidth]{figures/hmf.pdf}
% 	\caption{The cumulative MF (from \citealt{Tinker2008}) of halos above $M_{200c}$ at $z=0.1$}
% 	\label{fig: hmf}
% \end{figure}

@@

\subsection{Conditional {\rm[\ion{O}{II}]} Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS Data Release 12 \citep{Alam2015} catalogs to assign [\ion{O}{II}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{II}] line flux signal-to-noise of at least five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies, and the following SQL query selects the appropriate galaxies:
\begin{lstlisting}[language=SQL]
SELECT GAL.objid, 
       GAL.u, 
       GAL.err_u, 
       GAL.g, 
       GAl.err_g, 
       GAL.r, 
       GAL.err_r, 
       GAL.i, 
       GAl.err_i, 
       GAL.z, 
       GAL.err_z, 
       g.oii_3726_flux, 
       g.oii_3726_flux_err, 
       s.z as redshift 
FROM   galaxy AS GAL 
       JOIN specobj AS s 
         ON s.bestobjid = GAL.objid 
       JOIN galspecline AS g 
         ON g.specobjid = s.specobjid 
WHERE  oii_3726_flux > oii_3726_flux_err * 5 
       AND oii_3726_flux_err > 0 
       AND s.class = 'GALAXY' 
       AND s.zwarning = 0 
\end{lstlisting}

@@

The Hobby Eberly Telescope Dark Energy eXperiment (HETDEX; \citealt{Hill2008}) is a blind spectroscopic survey of 420 \degsq\ (300 \degsq, Spring field and 120 \degsq, Fall field) of the northern sky. The Spring field is $42\times7$ degrees centered at 13$^h\, +53\degree$, and the Fall field is $28\times5$ degrees centered at $1.5^h\, +0\degree$. HETDEX will observe high-redshift large scale structures using cutting edge wide-field integral field unit (IFU) spectrographs. The survey is designed to probe the evolution of the dark energy equation of state etched onto high redshift ($z>2$) galaxies by the Baryon Acoustic Oscillations (BAO) \citep{Eisenstein2005}. Tuned to find \lya\ emitting (LAE) galaxies at $1.9<z<3.5$, HETDEX expects to find 800,000 LAEs, and more than one million [\ion{O}{II}] emitting galaxies at $z<0.5$ masquerading as high-redshift galaxies \citep{Acquaviva2014}.

The Visible Integral-Field Replicable Unit Spectrograph (VIRUS; \citealt{Hill2012}) employed by HETDEX consists of 78 wide-field IFU spectrographs covering the wavelength region $3500 - 5500$~\AAA\ at R $\sim$ 750 \citep{Hill2008}. The spectral coverage allows for the detection of [\ion{O}{II}] ($\lambda\lambda 3727,3729$) emitters to $z\sim 0.5$ and Ca$^+$ H ($\lambda 3968.5$) and K ($\lambda 3933.7$) absorption features to $z\sim 0.4$. The $10 \sigma$ detection threshold for emission spectral features will be $3.5\times10^{-17}$ \ergscm\ at 5000 \AAA, or equivalently, $g = 21.9$ mag for continuum objects.

@@

The primary science goal of these observations is to provide $<1\%$ accuracy measurements of the Hubble expansion parameter and the angular diameter distance at $z\sim2$. This result will provide significant constraints on the evolution of the dark energy equation of state which is both competitive with, and independent of, constraints derived from observations of the \lya forest.

The objects of interest in this work are the low-$z$ [\ion{O}{II}] emitting galaxies, and while a large portion of the $\sim10^6$ interloping galaxies will be field (not associated with a bound structure) galaxies, the large area covered by HETDEX is expected to contain as many as 50 Virgo-sized (halo mass $>10^{15}$ \msol) clusters at $z<0.5$. The uniform spectroscopic coverage allows an unprecedentedly detailed look at a very large number of clusters ranging from group scales to the very massive. In addition to the recovery of accurate dynamical masses, detailed investigations of the dynamical state of the clusters becomes possible. However, it is unclear how a blind spectroscopic survey with an IFU will effect the recovery of galaxy cluster dynamical properties. Unlike many previous large cluster surveys \citeeg{Milvang-Jensen2008, Robotham2011, Sifon2015} which use multi-object spectrographs, the IFU spectrographs deployed by HETDEX sample the sky in a uniform but sparse pattern which could exclude member galaxies which would otherwise be included. Secondly, it is not straightforward to use spectroscopic redshifts predominately from emission-line galaxies to interpret the kinematic and dynamical states of the clusters. This work seeks to investigate both of these potential issues.
